Towards Comparable Active Learning

### 2023-05-31 doc/al_benchmark_paper.tex
divergingsin exp

\documentclass[]{article}
%\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
\usepackage{color}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{float}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{neurips_data_2023}

@@

\begin{abstract}
	Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
	In this paper we address the issue of inconsistent results in active learning literature.
	Authors of previous papers are constantly reporting significant performance improvements in their original work, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	We propose the first AL benchmark that tests algorithms in all three major domains of tabular, image and text data.
	Furthermore, we discuss overlooked problems for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment and provide evidence for the correct amount of repetitions for AL algorithms that reliably converge to the true median performance.
	We report empirical results for 6 algorithms on 7 datasets and aggregate them into a ranking of AL algorithms via Critical-Difference Diagrams.
\end{abstract}

\section{Introduction}
{\color{red} Version: 1.0}\\
Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
A problem that concerns academic researchers and practitioners in businesses alike and even could be extended to education in schools and hobbyists around the world.
On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
In their supplementary materials they conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while all following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm.
The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \# Datasets & Domains & Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 9 & 4 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem Description}
{\color{red} Version: 1.0}\\
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
	\quad\quad \text{\tiny (equivariant in the second argument)}
\end{align*}
% which is equivariant in the second argument, 
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal:
\begin{align*}
	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
	% \operatorname{avg} \{
	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
	\ell(y, \hat y(x)) 
	\\
	\text{with }
	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
	\\ 
	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
	\quad b\in 1{:}B
\end{align*}
(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Since combinatorial problem of finding the optimal subset $\mathcal{D}_\text{train}$ is computational not feasible, we allow sequential construction of $\mathcal{D}_\text{train}$. \\
%
To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$. \\
%
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{auc}(\mathcal{U}, \mathcal{L}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(y_{test}, \hat y_i(x_{test}))
\end{equation}
Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected.

\begin{minipage}{0.4\linewidth}
	\begin{algorithm}[H]
		\caption{Oracle}\label{alg:oracle}
		\begin{algorithmic}[1]
			\Ensure Retrain, margin
			\Require $\mathcal{U}^{(i)}, \mathcal{L}^{(i)}, \tau, \hat y_\theta$ 
			\State $\text{acc}^{(i)} \gets \operatorname{Retrain}(\mathcal{L}^{(i)})$ 
			\State $u^{(i)} \underset{\tau}{\sim} \text{unif}(1:|\mathcal{U}^{(i)}|)$
			\State $r^* \gets 0$
			\For{$t := 1 \ldots \tau$}
			\State $\mathcal{K} \gets \mathcal{L}^{(i)} \cup \{(u^{(i)}_t, y_t)\}$
			\State $\text{acc} \gets \operatorname{Retrain}(\mathcal{K}, \hat y_\theta)$  
			\State $r \gets \text{acc} - \text{acc}^{(i)}$
			\If{$r > r^*$} 
			\State $r^* \gets r$
			\State $t^* \gets t$
			\EndIf  
			\If{$r^* = 0$}
			\State $t^* \gets \operatorname{margin}(u^{(i)}, \hat y_\theta)$
			\EndIf
			\EndFor
			\Return $t^*$
		\end{algorithmic}
	\end{algorithm}
\end{minipage}
\hspace{3mm}
\begin{minipage}{0.59\linewidth}
	In Alg. \ref{alg:oracle} $\operatorname{Retrain}(\mathcal{L}^{(i)}, \hat y_\theta)$ trains the classification model $\hat y_\theta$ and returns the accuracy on the test set $\mathcal{D}_\text{test}$. \\ 
	$y_t$ is shorthand for the corresponding label of $u^{(i)}_{t}$ that can be recovered from the dataset labels. \\
	When the oracle does not find a sample with positive change in classification performance ($r^* = 0$), it applies margin sampling as a fallback ($\operatorname{margin}(u^{(i)}, \hat y_\theta)$).
\end{minipage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as newer approaches that have demonstrated significant lifts in their initial experiments.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches include CoreSet \cite{sener2017active} and TypiClust \cite{hacohen2022active}, which use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
Uncertainty-based approaches include classic uncertainty sampling (based on Shannon-Entropy and the margin-score), BALD \cite{kirsch2019batchbald} and BADGE \cite{ashdeep}, which use metrics to measure the classifiers state. \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations for data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
{\color{red} Version: 1.0}\\
We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split to mimic the leave-one-out protocol for cross-validation.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used all our 50 runs from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One AUC value like this corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can clearly observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To combat is uncertain evaluation, we propose to repeat every experiment 50 times, to arrive at the true median AUC for each algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reproducibility}\label{sec:reproducibility}
{\color{red} Version: 1.0}\\
A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding.
The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
This additional random number might be drawn during the initialization of the classification model or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
%This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
The desired control only be archived by assigning a separate random number generator to all these processes.
To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Oracle}\label{sec:oracle}
{\color{red} Version: 1.0}\\
Posing active learning as a sequence ordering problem, the oracle sequence for a given combination of dataset, model and training procedure would be the sequence that induces the highest AUC score for a given budget.
However, since this combinatorial problem is not solvable for realistic datasets, previous works have proposed approximations to this oracle sequence.
\cite{zhou2021towards} has used simulated annealing to search for the optimal sequence and used the best  solution found after a fixed time budget. 
Even though their reported performance curves display a significant lift over all other algorithms, we found the computational cost of reproducing this oracle for all our datasets to be prohibitive (The authors reported the search to take several days per dataset on 8 V100 GPUs). \\
In this paper we propose a greedy oracle algorithm that constructs an approximation of the optimal sequence in an iterative fashion.
Our oracle simply tests every data point in the provided sample of unlabeled points by fitting the classifier and directly measuring the resulting validation performance.
The point with the best validation performance is selected and added to the labeled pool for that iteration.
We noticed that this oracle is overfitting on the validation set, resulting in stagnating or even decreasing performance curves in later AL iterations.
To circumvent this problem, we introduced margin sampling as a fallback option for the oracle.
Whenever the oracle does not find an unlabeled point that results in an increase in performance (indicating an overfitting position), it defaults to margin sampling in that iteration.
The pseudocode for our oracle can be found in Alg. \ref{alg:oracle}.

>>>>

\documentclass[]{article}
%\usepackage[a4paper, total={6.5in, 8.5in}]{geometry}
\usepackage{color}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{neurips_data_2023}

@@

\begin{abstract}
	%Active Learning (AL) is a technique used in machine learning to improve data efficiency by selecting the most informative samples for annotation. 
	In this paper we address the issue of inconsistent results in active learning literature.
	Authors of previous papers are constantly reporting significant performance improvements, while subsequent literature fails to reproduce those results. This inconsistency leads to a chaotic landscape of AL algorithms.
	We propose the first AL benchmark that tests algorithms in three major domains of tabular, image and text data.
	Furthermore, we discuss overlooked problems for reproducing AL experiments with the default seeding setup that depends on a single seed per experiment and provide evidence for the correct amount of repetitions for AL algorithms that reliably converge to the true median performance.
	We report empirical results for 6 algorithms on 7 datasets and aggregate them into a ranking of AL algorithms via Critical-Difference Diagrams.
\end{abstract}

\section{Introduction}
Active Learning (AL) plays an important role in our society that applies machine learning to more and more areas and therefore has a high demand for labeled data in more and more areas.
A problem that concerns academic researchers and practitioners in businesses alike and even could be extended to education in schools and hobbyists around the world.
On top of providing a principled way to labeled unlabeled datasets, active learning is one of the two major approaches besides semi-supervised learning to make deep learning models more data efficient by requiring only a limited set of manually labeled data.
Both approaches are at their core orthogonal and can freely be combined and therefore we should continue our research efforts for both approaches. \\
Among others, the authors of \cite{zhou2021towards} have pointed out severe inconsistencies in results of AL papers in recent years. 
In their supplementary materials they conducted a meta analysis of reported results of several different AL algorithms and found that all considered algorithms only provided significant lifts in their own original papers, while all following literature reported performances no better that uncertainty sampling, or in some cases no better than random sampling for the same algorithm.
The result of these inconsistencies is a chaotic landscape of AL algorithms where every paper claims to archive state-of-the-art results by significantly outperforming everyone else, while the vast majority of results proves to be non-reproducible.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview}
We constrain our work on pool-based active learning where a pool of unlabeled samples is fixed at the start of each experiment and samples are chosen sequentially.
Specifically, we are not experimenting on so-called batch active learning, where at each iteration multiple unlabeled samples are chosen at the same time.
Even though batch AL is the more active research domain, it does not have a principled advantage over single-sample AL except speed of computation.
Not only is the problem of optimizing a portfolio of unlabeled samples more complicated to solve, the algorithms also have systematically less information per sample to work with.
For this reason we propose to focus more research effort on single-sample AL to find better algorithms in an environment that is simpler to solve and easier to control.
A performance comparison of batch AL and single-sample AL can be found in Fig. \ref{fig:ablation_bald}, which reproduces the message of Figure 1 from the paper \cite{gal2017deep} that proposed BALD, one of the SOTA algorithms for AL.
We can see that the batched version of BALD \cite{krizhevsky2009learning} can at most perform on-par with the single-sample algorithm.
Fig. \ref{fig:ablation_bald} also serves as a proof of concept for our provided code base.
Table \ref{tab:benchmark_comparison} shows a feature comparison between our proposed benchmark and several existing benchmarks in the literature, as well as methodological AL papers with experiments on at least two data domains.
\begin{table}[h]
	\centering
	\begin{tabular}{l | c c c c c}
		Paper & Sampling & \# Datasets & Domains & Algorithms & Oracle \\
		\hline
		Beck et al. \cite{beck2021effective} & batch & 4 & 1 & 7 & - \\
		Hu et al. \cite{hu2021towards} & batch & 5 & 2 & 13 & - \\
		Li et al. \cite{li2022empirical} & batch & 5 & 1 & 13 & - \\
		Zhou et al. \cite{zhou2021towards} & batch & 3 & 2 & 2 & \checkmark \\
		\textbf{Ours} & single & 9 & 4 & 6 & \checkmark 
	\end{tabular}
	\caption{Comparison of our benchmark with the existing literature}
	\label{tab:benchmark_comparison}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Problem Description}
Given two spaces $\mathcal{X}:=\mathbb{R}^M$ and $\mathcal{Y}:=\mathbb{R}^C$ where a samples is drawn from an unknown data distribution $(x, y) \sim p$ with $x \in \mathcal{X}, y \in \mathcal{Y}$.
We call $\mathcal{L} := \{ (x, y) \}^*$ the labeled pool and $\mathcal{U} := \{ x \}^*$ the unlabeled pool.
Their domains are $\Omega := \mathbb{R}^M \times \mathbb{R}^C$ and $\Lambda := \mathbb{R}^M$ respectively. \\
A function $\hat y: \mathcal{X} \rightarrow \mathcal{Y}$ is called a classifier and a function $\ell: \mathcal{Y} \rightarrow \mathcal{Y}$ is called a loss.
Given a scalar $B := \mathbb{R}; \hspace{1mm} B < |\mathcal{U}|$ called budget, we want to find an acquisition function $a: \Omega \times \Lambda \mapsto \mathcal{U}^B$ that creates a subset of $\mathcal{U}$ so that the expected loss on our data is minimal.
$$
\underset{(x,y)\sim p}{\mathbb{E}} \ell(y, \hat y(x))
$$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lars' Problem Description}
Given two spaces $\mathcal{X}:=\mathcal{R}^M$ and $\mathcal{Y}:=\mathcal{R}^C$, a sample $\mathcal{D}_1,\ldots,\mathcal{D}_N \subseteq (\mathcal{X}\times \mathcal{Y})^*$ of sequences of pairs $(x,y)$  from an unknown distribution $p$ called datasets and a number $B\in\mathcal{N}$ with $B < |\mathcal{D}|$. \\
Given two functions $\ell:\mathcal{Y}\times \mathcal{Y}\rightarrow \mathcal{R}$ called loss, and $A: (\mathcal{X} \times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \mathcal{Y}^{\mathcal{X}}$ called learning algorithm, find a function
\begin{align*}
	a: (\mathcal{X}\times \mathcal{Y})^* \times \mathcal{X}^* \rightarrow \{0,1\}^*
\end{align*}
called acquisition function, s.t. the expected loss of a model learned on all predictors plus $B$ acquired targets is minimal.
\begin{align*}
	\mathbb{E}_{\mathcal{D}\text{train},\mathcal{D}\text{test}\sim p}   &
	% \frac{1}{|\mathcal{D}\text{test}|} \sum_{(x,y)\in\mathcal{D}\text{test}}
	% \operatorname{avg} \{
	%  \ell(y, \hat y(x)) \mid (x,y)\in\mathcal{D}\text{test} \}
	\operatorname{avg}\limits_{(x,y)\in\mathcal{D}\text{test}}
	\ell(y, \hat y(x)) 
	\\
	\text{with }
	\hat y:= & A( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_B}), \mathcal{D}\text{train}|_{\mathcal{X}})
	\\ 
	n_b := & \text{index}( a( (\mathcal{D}\text{train}_{n_1},\ldots,\mathcal{D}\text{train}_{n_{b-1}}), \mathcal{D}\text{train}|_{\mathcal{X}}) ),
	\quad b\in 1{:}B
\end{align*}
%(\textit{We would need to switch from lowest expected loss to highest AUC}) \\
Even though the acquisition function in principle could output the full subset of $\mathcal{D}_\text{train}$, the combinatorial problem is computationally not feasible and we allow sequential construction of the subset as a relaxation of the problem. \\
%
To construct the active learning setting from existing labeled datasets, we first split it into $\mathcal{D}_\text{train}$, $\mathcal{D}_\text{val}$ and $\mathcal{D}_\text{test}$ and then suppress the labels $\mathcal{Y}$ of $\mathcal{D}_\text{train}$ to form the unlabeled pool $\mathcal{U} := \mathcal{X}^*$ and form and initial labeled pool by uniformly sampling $k$ number of instances per class from $\mathcal{U}$ and recovering their label $\mathcal{L} := (\mathcal{X}, \mathcal{Y})^{k*C}$. \\
%
Following \cite{zhou2021towards}, the quality of an active learning algorithm is evaluated by an "anytime" protocol that incorporates classification performance at every iteration, not just the final performance after the budget is exhausted.
We employ the normalized area under the accuracy curve (AUC):
\begin{equation}\label{eq:auc}
	\operatorname{auc}(\mathcal{D}_{test}, \hat y, B) := \frac{1}{B} \sum_{i=1}^{B} \operatorname{Acc}(\mathcal{D}_{test}, \hat y_i)
\end{equation}
Where $\hat y_i$ is the retrained classification model after the i-\textit{th} instance was selected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
{\color{red} Version: Braindump}\\
Many different algorithms have been proposed for active learning. 
In this work we focus on those approaches that have shown consistent results over the years as well as some of the new approaches.
AL algorithms can be categorized into two classes: Geometric approaches and uncertainty-based approaches.
Geometric approaches use clustering techniques to partition the data and then sample their unlabeled points based on the clusters.
They often use the current classification model $\hat y_i$ to encode the data into a latent space to improve the performance of their clustering.
This benchmark includes the following geometric approaches: CoreSet \cite{sener2017active}, BADGE \cite{ashdeep} and TypiClust \cite{hacohen2022active}.
Uncertainty-based approaches use metrics to measure the classifiers state.
Commonly, a proxy for the sought after uncertainty of the model for a given datapoint is the distance of that point to the various decision boundaries, measured via the softmax output of the model.
This benchmark includes Shannon-Entropy sampling \cite{wang2014new}, margin sampling \cite{wang2014new} and BALD \cite{kirsch2019batchbald} \\ [1mm]
%
Some previous work also aimed to provide a benchmark suite for active learning:
The authors of \cite{beck2021effective} and \cite{li2022empirical} both focus on active learning in the image domain.
While \cite{beck2021effective} discuss a new metric to measure AL performance, which they call "Label Efficiency" and provide experiments on many common configurations of data preparation, model training and other hyperparameters, \cite{li2022empirical} focuses on combined approaches of AL and semi-supervised learning to aid model training.
The authors of \cite{hu2021towards} study models that are learned with AL techniques in the image and text domain.
They test for several different properties of the models including robustness, response to compression techniques and final performance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}\label{sec:evaluation}
We compare different AL algorithms based on their median AUC score (Eq. \ref{eq:auc}) across multiple restarts of the experiment.
Each restart will retain the train/test split (often given by the dataset itself), but introduces a new validation split to mimic the leave-one-out protocol for cross-validation.
The AUC incorporates performance in early stages (low budget) as well as capabilities to push the classifier in later stages (high budget).
A good AL algorithm should be able to perform well in both scenarios. \\ [1mm]
Since AUC is dependent on the chosen budget, we need a general rule on how to set this hyperparameter that does not inherently benefit a subset of algorithms.
In this work, we choose the budget per dataset to be the first point at which any algorithm (except oracle) manages to reach a percentage of the upper bound performance measured on the full dataset.
Even though we would like to propose a single percentage value for all datasets, we found that different data modalities and use cases need different percentages to produce sensible budgets.
We propose the following values: \textbf{Tabular}: 99\%, \textbf{Image}: 90\% and \textbf{Text}: 95\%. \\ [1mm]
Additionally, we provide evidence in Fig. \ref{fig:restarts} that previous works have not evaluated their experiments with a sufficient number of restarts.
To create Fig. \ref{fig:restarts} we used all our 50 runs from the margin/random sampling algorithm on the splice dataset.
From these 50 runs we uniformly sampled subsets of runs and calculated the median AUC for this subset.
One AUC value like this corresponds to one cross-validated experiment sampled from the distribution of experiments that are restarted exactly this many times.
To create one slice in Fig. \ref{fig:restarts}, we drew 50 samples from this distribution.
Each box-plot represents the variance of an evaluation if conducted with the respective number of restarts.
We can clearly observe that low repetitions ($<10$) provide an uncertain evaluation where lucky and unlucky draws of the same experiment give drastically different median AUC values.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{img/ablation_restarts.png}
	\caption{Random draws from an experimental distribution on the Splice datasets with different numbers of repetitions. Each point on the Y-axis represents a cross-validated result that could have been reported in a paper. This analysis shows the drastic differences in performance one could observe even when repeating an experiment 2-10 times.}
	\label{fig:restarts}
\end{figure}
To combat is uncertain evaluation, we propose to repeat every experiment 50 times, to arrive at the true median AUC for each algorithm.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reproducibility}\label{sec:reproducibility}
A big focus in this work is to provide an experimental setup that is fully reproducible independent of the dataset, classification model or AL algorithm used.
Given a seed, an evaluation on one dataset should always be done on the same validation split as well as the same random state for all included systems, like the mini batch sampler for model training or the initialization for the classifier itself.
Even though different AL algorithms will pick different samples, making them unavailable for sampling in earlier or later batches, the theoretical decision tree for every possible choice in every iteration $i$ should stay the same.
Since every possible trajectory cannot be precomputed and stored to disk, we need to resort to seeding.
The default choice of setting a global seed at the start of the experiment is not sufficient here, since a single additional random draw from the random number generator completely changes the behavior of all other systems.
This additional random number might be drawn during the initialization of the classification model or the AL algorithm, or even during every AL iteration if $\Omega$ is stochastic.
%This problem also applies to the initialization of the classification model $\theta$ and the initialization and querying of $\Omega$.
The desired control only be archived by assigning a separate random number generator to all these processes.
To the best of our knowledge, we are the first work that discusses this issue and proposes a solution for it.
We hypothesize that the insufficient setup with global seeds contributes to the on-going problem of inconsistent results of AL algorithms in different papers. \\ [1mm]
In summary, we introduce three different seeds: $s_\Omega$ for the acquisition function, $s_\mathcal{D}$ for dataset splitting and mini batch sampling and $s_\theta$ for model initialization and sampling of dropout masks.
Unless stated otherwise, we will keep $s_\Omega$ fixed for restarts of the same experiment, while $s_\mathcal{D}$ and $s_\theta$ are incremented by 1 between restarts to introduce stochasticity into our framework.
% This seeding setup enables a reproduction of the oracle curve even when the classifier or AL algorithm is different.
